---
title: "Binary Regression"
author: "Ramin Mojab"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Seemingly Unrelated Regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "  >"
)
```

```{r}
library(ldt) 
set.seed(123)
```


## Introduction
In this vignette, I will introduce you to the `ldt` package and its main features for dealing with Logistic and Probit Regression models. You will learn how to estimate a binary model, make predictions, and assess model uncertainty. Additionally, we will explore the use of Principal Component Analysis as an alternative approach for handling a large number of potential explanatory variables.

One of the key ideas behind `ldt` is to minimize user discretion by using a rule-based approach to select data. This approach not only avoids discretion but also automates the process of searching for the best models within a defined model set.

To demonstrate these features, I will create an artificial dataset with a dependent variable and both relevant and irrelevant explanatory variables. The dependent and relevant explanatory variables are sampled from a known binary model. While we can evaluate how well the estimation process finds the true parameters, our main focus will be on how to estimate, search, predict, and report results.

Let’s get started!



## A simple experiment
Let’s start by assuming that we know the structure of the model. We can do this by simulating data from two known binary models. The following command generates the required samples:

```{r}
num_obs <- 100 
num_exo <- 3L 
p_positive <- 0.4
sample_l <- sim.dc(num_exo, num_obs, probit = FALSE, pPos = p_positive)
sample_p <- sim.dc(sample1$coef, num_obs, probit = TRUE, pPos = p_positive)

print(sample1$coef)
```

We know the parameters of the systems because they are included in the output of the `sim.dc` function. Note that there is a logit and a probit model. Each model has one equation or dependent variable. This equation has an intercept and `r num_exo-1` exogenous variables. The sample size is `r num_obs`. The coefficient vector of the logit model is generated randomly and is listed in the output, `sample_l`. It is used to sample data from the logit model and therefore, the parameters of the two models are the same. In these samples, the `r p_positive*100` percent of observations are labeled as positive (because of `pPos` argument).

The LaTeX code for the equations of the system is in the `eqLatex` element. It results in the following representation:




